Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-03-04 article-models-tt.tex
code cleanup, plus fixes of small bugs

\begin{defn}
When $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$ are terms,
we define 
\begin{equation*}
\jtermdefn{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}{\subst{u}{{x}{\idfunc[\ctxext{A}{P}]}}}.
\end{equation*} 
\end{defn}

Thus, the term $\ctxext{x}{u}$ is the pairing of $x$ and $u$. Note that because
we have the judgmental equality 
$\ctxwk{P}{{A}{\ctxext{A}{P}}}\jdeq\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}$ in the
context $\ctxext{{\Gamma}{A}}{P}$, the
pairing function could just be defined as $\idfunc[\ctxext{A}{P}]$.

@@

\begin{lem}
The following inference rules are valid:
\begin{align*}
& \inference
  {\jterm{\Gamma}{A}{x}
   \jterm{\Gamma}{\subst{x}{P}}{u}
   \jfam{{{\Gamma}{A}}{P}}{Q}}
  {\jfameq{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{\subst{u}{{x}{Q}}}}\\
& \inference
  {\jterm{\Gamma}{A}{x}
   \jterm{\Gamma}{\subst{x}{P}}{u}
   \jterm{{{\Gamma}{A}}{P}}{Q}{g}}
  {\jtermeq{\Gamma}{\subst{u}{{x}{Q}}}{\subst{\ctxext{x}{u}}{g}}{\subst{u}{{x}{g}}}}\\
& \inference
  {\jterm{\Gamma}{A}{x}
   \jterm{\Gamma}{\subst{x}{P}}{u}
   \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}}
  {\jfameq{{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{\ctxext{x}{u}}{R}}{\subst{u}{{x}{R}}}}\\
& \inference
  {\jterm{\Gamma}{A}{x}
   \jterm{\Gamma}{\subst{x}{P}}{u}
   \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{t}}
  {\jtermeq{{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{u}{{x}{R}}}{\subst{\ctxext{x}{u}}{t}}{\subst{u}{{x}{t}}}}
\end{align*}
\end{lem}

\begin{proof}
We prove only the first judgmental equality. All the others are similar.
Let $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$
be terms and let $\jfam{{{\Gamma}{A}}{P}}{Q}$ be a family. Then we have
\begin{align*}
\subst{\ctxext{x}{u}}{Q} 
& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{Q} \tag{by definition}\\
& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{{x}{\ctxwk{A}{Q}}}\tag{by \autoref{defn-ws-3}}\\
& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\ctxwk{A}{Q}}}}}\tag{by \autoref{defn-ws-3}}\\
& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{{u}{{x}{\ctxwk{P}{{A}{Q}}}}}\tag{by \autoref{comp-sw-f}}\\
& \jdeq \subst{u}{{{x}{\idfunc[\ctxext{A}{P}]}}{{x}{\ctxwk{{P}{{A}{Q}}}}}}\tag{by \autoref{comp-ss-f}}\\
& \jdeq \subst{u}{{x}{{\idfunc[\ctxext{A}{P}]}{\ctxwk{P}{{A}{Q}}}}}\tag{by \autoref{comp-ss-f}}\\
& \jdeq \subst{u}{{x}{{\idfunc[\ctxext{A}{P}]}{\ctxwk{\ctxext{A}{P}}{Q}}}}\tag{by \autoref{comp-ew-f}}\\
& \jdeq \subst{u}{{x}{Q}}\tag{by \autoref{idfunc-wk-defn}}
\end{align*}
\end{proof}

@@

\begin{defn}
Let $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$ be families. We define
\begin{align*}
\jtermdefn*{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{P}{{A}{A}}}{\cprojfstf{A}{P}}{\ctxwk{P}{\idfunc[A]}}\\
\jtermdefn*{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{P}{P}}{\cprojsndf{A}{P}}{\idfunc[P]}
\end{align*}
\end{defn}

@@

\begin{lem}
For any $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$ we have
\begin{equation}\label{idfunc-ext-comp}
\inference
  {\jfam{\Gamma}{A}
   \jfam{{\Gamma}{A}}{P}}
  {\jtermeq
    {{\Gamma}{{A}{P}}}
    {\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}
    {\idfunc[\ctxext{A}{P}]}
    {\ctxext{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}}}
\end{equation}
\end{lem}

\begin{proof}
Consider the families $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$. Then
we have the judgmental equalities
\begin{align*}
\ctxext{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}
& \jdeq \subst{\cprojsndf{A}{P}}{{\cprojfstf{A}{P}}{\idfunc[\ctxext{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{\ctxext{A}{P}}{P}}]}}\\
& \jdeq \subst
          {\cprojsndf{A}{P}}
          { {\cprojfstf{A}{P}}
            {\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}\\
& \jdeq \subst
          {\idfunc[P]}
          { {\ctxwk{P}{\idfunc[A]}}
            {\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}\\
& \jdeq \subst
          {\idfunc[P]}
          { {\ctxwk{P}{\idfunc[A]}}
            {\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}}\\
& \jdeq \subst
          {\idfunc[P]}
          { {\ctxwk{P}{\idfunc[A]}}
            {\ctxwk{P}{{A}{\idfunc[\ctxext{A}{P}]}}}}\\
& \jdeq \subst
          {\idfunc[P]}
          {\ctxwk
            {P}
            {\subst
              {\idfunc[A]}
              {\ctxwk{A}{\idfunc[\ctxext{A}{P}]}}}}\\
& \jdeq \subst
          {\idfunc[A]}
          {\ctxwk{A}{\idfunc[\ctxext{A}{P}]}}\\
& \jdeq \idfunc[\ctxext{A}{P}].
\end{align*}
\end{proof}

@@

The constructions of the terms $\ctxext{x}{u}$ and $\cprojfst{A}{P}{w}$ and
$\cprojsnd{A}{P}{w}$ are subject to various rules, with all of them being
consequences of earlier introduced inference rules.

\begin{lem} The following inference rules expressing that pairing is a strict
inverse to the combination of decompositions, are valid:
\begin{align*}
& \inference
  {\jterm{\Gamma}{\ctxext{A}{P}}{w}}
  {\jtermeq{\Gamma}{\ctxext{A}{P}}{\ctxext{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}}{w}}\\
& \inference
  {\jterm{\Gamma}{A}{x}
   \jterm{\Gamma}{\subst{x}{P}}{u}}
  {\jtermeq{\Gamma}{A}{\cprojfst{A}{P}{\ctxext{x}{u}}}{x}}\\
& \inference
  {\jterm{\Gamma}{A}{x}
   \jterm{\Gamma}{\subst{x}{P}}{u}}
  {\jtermeq{\Gamma}{\subst{x}{P}}{\cprojsnd{A}{P}{\ctxext{x}{u}}}{u}}
\end{align*}
\end{lem}

\begin{proof}
To prove the first judgmental equality, note that
\begin{align*}
w & \jdeq \subst{w}{\idfunc[\ctxext{A}{P}]} \tag{by \autoref{idfunc-subst-defn}}\\
& \jdeq \subst{w}{\subst{\idfunc[P]}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}
  \tag{by \autoref{idfunc-ext-comp}}\\
& \jdeq \subst{{w}{\idfunc[P]}}{\subst{w}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}\tag{by \autoref{comp-ss-t}}\\
& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\subst{w}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}\tag{by \autoref{comp-ss-t}}\\
& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\subst{w}{\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}}}\tag{by \autoref{idfunc-wk-comp}}\\
& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\idfunc[\ctxext{A}{P}]}}\tag{by \autoref{defn-ws-4}}\\
& \jdeq \ctxext{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}\tag{by definition}
\end{align*}
To prove the second judgmental equality, let $\jterm{\Gamma}{A}{x}$ and
$\jterm{\Gamma}{\subst{x}{P}}{u}$. Then we have
\begin{align*}
\cprojfst{A}{P}{\ctxext{x}{u}}
& \jdeq \subst{\ctxext{x}{u}}{\ctxwk{P}{\idfunc[A]}}\\
& \jdeq \subst{u}{{x}{\ctxwk{P}{\idfunc[A]}}} \\
& \jdeq \subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\idfunc[A]}}}\\
& \jdeq \subst{x}{\idfunc[A]}\\
& \jdeq x.
\end{align*}
To prove the third judgmental equality, note that
\begin{align*}
\cprojsnd{A}{P}{\ctxext{x}{u}}
& \jdeq \subst{\ctxext{x}{u}}{\idfunc[P]}\\
& \jdeq \subst{u}{{x}{\idfunc[P]}}\\
& \jdeq \subst{u}{\idfunc[\subst{x}{P}]}\\
& \jdeq u.
\end{align*}
\end{proof}

\begin{lem}
The following compatibility rules for two consecutive term extensions is valid:
\begin{align*}
& \inference
  {\jterm{\Gamma}{A}{x}
   \jterm{\Gamma}{\subst{x}{P}}{u}
   \jterm{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{v}}
  {\jtermeq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{x}{{u}{v}}}{\ctxext{{x}{u}}{v}}}\\
& \inference
  {\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
  {\jtermeq{\Gamma}{A}{\cprojfst{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}}}{\cprojfst{A}{\ctxext{P}{Q}}{w}}}\\
& \inference
  {\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
  {\jtermeq{\Gamma}{\subst{\cprojfst{A}{\ctxext{P}{Q}}{w}}{P}}
    {\cprojsnd{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}}}
    {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}}\\
& \inference
  {\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
  {\jtermeq{\Gamma}{\subst{\cprojfst{\ctxext{A}{P}}{Q}{w}}{Q}}
    {\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
    {\cprojsnd{\ctxext{A}{P}}{Q}{w}}}
\end{align*}
\end{lem}

\begin{proof}
Consider terms $\jterm{\Gamma}{A}{x}$, $\jterm{\Gamma}{\subst{x}{P}}{u}$ and
$\jterm{\Gamma}{\subst{u}{{x}{Q}}}{v}$. Then we have
\begin{align*}
\ctxext{x}{{u}{v}} & \jdeq \subst{\ctxext{u}{v}}{{x}{\idfunc[\ctxext{A}{{P}{Q}}]}}\\
& \jdeq \subst{v}{{u}{{x}{\idfunc[\ctxext{A}{{P}{Q}}]}}}\\
& \jdeq \subst{v}{{u}{{x}{\idfunc[\ctxext{{A}{P}}{Q}]}}}\\
& \jdeq \subst{v}{{\ctxext{x}{u}}{\idfunc[\ctxext{{A}{P}}{Q}]}}\\
& \jdeq \ctxext{{x}{u}}{v}.
\end{align*}
Now consider a term $\jterm{\Gamma}{\ctxext{A}{{P}{Q}}}{w}$. Then we have
\begin{align*}
w 
& \jdeq \ctxext{\cprojfst{A}{\ctxext{P}{Q}}{w}}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}\\
& \jdeq \ctxext{\cprojfst{A}{\ctxext{P}{Q}}{w}}{{\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}{\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}}\\
& \jdeq \ctxext{{\cprojfst{A}{\ctxext{P}{Q}}{w}}{\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}}{\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
\end{align*}
Thus we see that 
\begin{align*}
\cprojfst{\ctxext{A}{P}}{Q}{w} & \jdeq \ctxext{\cprojfst{A}{\ctxext{P}{Q}}{w}}{\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}\\ 
\cprojsnd{\ctxext{A}{P}}{Q}{w} & \jdeq \cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
\end{align*}
proving the fourth judgmental equality, and therefore also that
\begin{align*}
\cprojfst{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}} & \jdeq \cprojfst{A}{\ctxext{P}{Q}}{w}\\
\cprojsnd{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}} & \jdeq \cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
\end{align*}
proving the second and the third judgmental equalities.
\end{proof}

\begin{lem}
When we weaken a term $\ctxext{y}{v}$ of $\ctxext{B}{Q}$ in context $\Gamma$ by
a family $A$, the term that we get is $\ctxext{\ctxwk{A}{y}}{\ctxwk{A}{v}}$. More
precisely, the following inference rules are valid:
\begin{align*}
& \inference
  {\jterm{{\Gamma}{B}}{Q}{g}
   \jterm{{\Gamma}{B}}{\subst{g}{R}}{t}}
  {\jtermeq
    {{{\Gamma}{A}}{\ctxwk{A}{B}}}
    {\ctxwk{A}{\ctxext{Q}{R}}}
    {\ctxwk{A}{\ctxext{g}{t}}}
    {\ctxext{\ctxwk{A}{g}}{\ctxwk{A}{t}}}}
% \\
%& \inference{\jterm{\Gamma}{\ctxwk{A}{\ctxext{B}{Q}}}{\ctxwk{A}{w}}}{\jtermeq{{\Gamma}{A}}{\ctxwk{A}{B}}{\cprojfst{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A}{w}}}{\ctxwk{A}{(\cprojfst{B}{Q}{w})}}}\\
\end{align*}
\end{lem}

>>>>

\begin{defn}
When $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$ are terms,
we define 
\begin{equation*}
\jtermdefn{\Gamma}{\ctxext{A}{P}}{\tmext{A}{P}{x}{u}}{\unfold{\tmext{A}{P}{x}{u}}}.
\end{equation*} 
\end{defn}

Thus, the term $\tmext{A}{P}{x}{u}$ is the pairing of $x$ and $u$. Note that because
we have the judgmental equality 
$\ctxwk{P}{{A}{\ctxext{A}{P}}}\jdeq\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}$ in the
context $\ctxext{{\Gamma}{A}}{P}$, the
pairing function could just be defined as $\idfunc[\ctxext{A}{P}]$.

@@

\begin{lem}
The following inference rules are valid:
\begin{align*}
& \inference
  {\jterm{\Gamma}{A}{x}
   \jterm{\Gamma}{\subst{x}{P}}{u}
   \jfam{{{\Gamma}{A}}{P}}{Q}}
  {\jfameq{\Gamma}{\subst{\tmext{A}{P}{x}{u}}{Q}}{\subst{u}{{x}{Q}}}}\\
& \inference
  {\jterm{\Gamma}{A}{x}
   \jterm{\Gamma}{\subst{x}{P}}{u}
   \jterm{{{\Gamma}{A}}{P}}{Q}{g}}
  {\jtermeq{\Gamma}{\subst{u}{{x}{Q}}}{\subst{\tmext{A}{P}{x}{u}}{g}}{\subst{u}{{x}{g}}}}\\
& \inference
  {\jterm{\Gamma}{A}{x}
   \jterm{\Gamma}{\subst{x}{P}}{u}
   \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}}
  {\jfameq{{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{\tmext{A}{P}{x}{u}}{R}}{\subst{u}{{x}{R}}}}\\
& \inference
  {\jterm{\Gamma}{A}{x}
   \jterm{\Gamma}{\subst{x}{P}}{u}
   \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{t}}
  {\jtermeq{{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{u}{{x}{R}}}{\subst{\tmext{A}{P}{x}{u}}{t}}{\subst{u}{{x}{t}}}}
\end{align*}
\end{lem}

\begin{proof}
We prove only the first judgmental equality. All the others are similar.
Let $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$
be terms and let $\jfam{{{\Gamma}{A}}{P}}{Q}$ be a family. Then we have
\begin{align*}
\subst{\tmext{A}{P}{x}{u}}{Q} 
& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{Q} \tag{by definition}\\
& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{{x}{\ctxwk{A}{Q}}}\tag{by \autoref{defn-ws-3}}\\
& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\ctxwk{A}{Q}}}}}\tag{by \autoref{defn-ws-3}}\\
& \jdeq \subst{{u}{{x}{\idfunc[\ctxext{A}{P}]}}}{{u}{{x}{\ctxwk{P}{{A}{Q}}}}}\tag{by \autoref{comp-sw-f}}\\
& \jdeq \subst{u}{{{x}{\idfunc[\ctxext{A}{P}]}}{{x}{\ctxwk{{P}{{A}{Q}}}}}}\tag{by \autoref{comp-ss-f}}\\
& \jdeq \subst{u}{{x}{{\idfunc[\ctxext{A}{P}]}{\ctxwk{P}{{A}{Q}}}}}\tag{by \autoref{comp-ss-f}}\\
& \jdeq \subst{u}{{x}{{\idfunc[\ctxext{A}{P}]}{\ctxwk{\ctxext{A}{P}}{Q}}}}\tag{by \autoref{comp-ew-f}}\\
& \jdeq \subst{u}{{x}{Q}}\tag{by \autoref{idfunc-wk-defn}}
\end{align*}
\end{proof}

@@

\begin{defn}
Let $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$ be families. We define
\begin{align*}
\jhomdefn*{\Gamma}{{A}{P}}{A}{\cprojfstf{A}{P}}{\unfold{\cprojfstf{A}{P}}}\\
\jtermdefn*{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{P}{P}}{\cprojsndf{A}{P}}{\idfunc[P]}
\end{align*}
\end{defn}

@@

\begin{lem}
For any $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$ we have
\begin{equation}\label{idfunc-ext-comp}
\inference
  {\jfam{\Gamma}{A}
   \jfam{{\Gamma}{A}}{P}}
  {\jhomeq{\Gamma}{{A}{P}}{{A}{P}}
    {\idfunc[\ctxext{A}{P}]}
    {\tmext{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{\ctxext{A}{P}}{P}}{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}}}
\end{equation}
\end{lem}

\begin{proof}
Consider the families $\jfam{\Gamma}{A}$ and $\jfam{{\Gamma}{A}}{P}$. Then
we have the judgmental equalities
\begin{align*}
\tmext{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{\ctxext{A}{P}}{P}}{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}
& \jdeq \unfold{\tmext{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{\ctxext{A}{P}}{P}}{\cprojfstf{A}{P}}{\cprojsndf{A}{P}}}\\
& \jdeq \subst
          {\idfunc[P]}
          { {\ctxwk{P}{\idfunc[A]}}
            {\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}\\
& \jdeq \subst
          {\idfunc[P]}
          { {\ctxwk{P}{\idfunc[A]}}
            {\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}}\\
& \jdeq \subst
          {\idfunc[P]}
          { {\ctxwk{P}{\idfunc[A]}}
            {\ctxwk{P}{{A}{\idfunc[\ctxext{A}{P}]}}}}\\
& \jdeq \subst
          {\idfunc[P]}
          {\ctxwk
            {P}
            {\subst
              {\idfunc[A]}
              {\ctxwk{A}{\idfunc[\ctxext{A}{P}]}}}}\\
& \jdeq \subst
          {\idfunc[A]}
          {\ctxwk{A}{\idfunc[\ctxext{A}{P}]}}\\
& \jdeq \idfunc[\ctxext{A}{P}].
\end{align*}
\end{proof}

@@

The constructions of the terms $\tmext{A}{P}{x}{u}$ and $\cprojfst{A}{P}{w}$ and
$\cprojsnd{A}{P}{w}$ are subject to various rules, with all of them being
consequences of earlier introduced inference rules.

\begin{lem} The following inference rules expressing that pairing is a strict
inverse to the combination of decompositions, are valid:
\begin{align*}
& \inference
    {\jterm{\Gamma}{\ctxext{A}{P}}{w}}
    {\jtermeq
      {\Gamma}
      {\ctxext{A}{P}}
      {\tmext{A}{P}{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}}
      {w}}\\
& \inference
  {\jterm{\Gamma}{A}{x}
   \jterm{\Gamma}{\subst{x}{P}}{u}}
  {\jtermeq{\Gamma}{A}{\cprojfst{A}{P}{\tmext{A}{P}{x}{u}}}{x}}\\
& \inference
  {\jterm{\Gamma}{A}{x}
   \jterm{\Gamma}{\subst{x}{P}}{u}}
  {\jtermeq{\Gamma}{\subst{x}{P}}{\cprojsnd{A}{P}{\tmext{A}{P}{x}{u}}}{u}}
\end{align*}
\end{lem}

\begin{proof}
To prove the first judgmental equality, note that
\begin{align*}
w & \jdeq \subst{w}{\idfunc[\ctxext{A}{P}]} \tag{by \autoref{idfunc-subst-defn}}\\
& \jdeq \subst{w}{\subst{\idfunc[P]}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}
  \tag{by \autoref{idfunc-ext-comp}}\\
& \jdeq \subst{{w}{\idfunc[P]}}{\subst{w}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}\tag{by \autoref{comp-ss-t}}\\
& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\subst{w}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}\tag{by \autoref{comp-ss-t}}\\
& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\subst{w}{\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}}}\tag{by \autoref{idfunc-wk-comp}}\\
& \jdeq \subst{{w}{\idfunc[P]}}{{\subst{w}{\ctxwk{P}{\idfunc[A]}}}{\idfunc[\ctxext{A}{P}]}}\tag{by \autoref{defn-ws-4}}\\
& \jdeq \tmext{A}{P}{\cprojfst{A}{P}{w}}{\cprojsnd{A}{P}{w}}\tag{by definition}
\end{align*}
To prove the second judgmental equality, let $\jterm{\Gamma}{A}{x}$ and
$\jterm{\Gamma}{\subst{x}{P}}{u}$. Then we have
\begin{align*}
\cprojfst{A}{P}{\tmext{A}{P}{x}{u}}
& \jdeq \subst{\tmext{A}{P}{x}{u}}{\ctxwk{P}{\idfunc[A]}}\\
& \jdeq \subst{u}{{x}{\ctxwk{P}{\idfunc[A]}}} \\
& \jdeq \subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\idfunc[A]}}}\\
& \jdeq \subst{x}{\idfunc[A]}\\
& \jdeq x.
\end{align*}
To prove the third judgmental equality, note that
\begin{align*}
\cprojsnd{A}{P}{\tmext{A}{P}{x}{u}}
& \jdeq \subst{\tmext{A}{P}{x}{u}}{\idfunc[P]}\\
& \jdeq \subst{u}{{x}{\idfunc[P]}}\\
& \jdeq \subst{u}{\idfunc[\subst{x}{P}]}\\
& \jdeq u.
\end{align*}
\end{proof}

\begin{lem}
The following compatibility rules for two consecutive term extensions is valid:
\begin{align*}
& \inference
  {\jterm{\Gamma}{A}{x}
   \jterm{\Gamma}{\subst{x}{P}}{u}
   \jterm{\Gamma}{\subst{\tmext{A}{P}{x}{u}}{Q}}{v}}
  {\jtermeq{\Gamma}{\ctxext{{A}{P}}{Q}}{\tmext{A}{{P}{Q}}{x}{{u}{v}}}{\tmext{{A}{P}}{Q}{{x}{u}}{v}}}\\
& \inference
  {\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
  {\jtermeq{\Gamma}{A}{\cprojfst{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}}}{\cprojfst{A}{\ctxext{P}{Q}}{w}}}\\
& \inference
  {\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
  {\jtermeq{\Gamma}{\subst{\cprojfst{A}{\ctxext{P}{Q}}{w}}{P}}
    {\cprojsnd{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}}}
    {\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}}\\
& \inference
  {\jterm{\Gamma}{\ctxext{{A}{P}}{Q}}{w}}
  {\jtermeq{\Gamma}{\subst{\cprojfst{\ctxext{A}{P}}{Q}{w}}{Q}}
    {\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
    {\cprojsnd{\ctxext{A}{P}}{Q}{w}}}
\end{align*}
\end{lem}

\begin{proof}
Consider terms $\jterm{\Gamma}{A}{x}$, $\jterm{\Gamma}{\subst{x}{P}}{u}$ and
$\jterm{\Gamma}{\subst{u}{{x}{Q}}}{v}$. Then we have
\begin{align*}
\tmext{A}{{P}{Q}}{x}{{u}{v}} & \jdeq \subst{\tmext{P}{Q}{u}{v}}{{x}{\idfunc[\ctxext{A}{{P}{Q}}]}}\\
& \jdeq \subst{v}{{u}{{x}{\idfunc[\ctxext{A}{{P}{Q}}]}}}\\
& \jdeq \subst{v}{{u}{{x}{\idfunc[\ctxext{{A}{P}}{Q}]}}}\\
& \jdeq \subst{v}{{\tmext{A}{P}{x}{u}}{\idfunc[\ctxext{{A}{P}}{Q}]}}\\
& \jdeq \tmext{{A}{P}}{Q}{{x}{u}}{v}.
\end{align*}
Now consider a term $\jterm{\Gamma}{\ctxext{A}{{P}{Q}}}{w}$. Then we have
\begin{align*}
w 
& \jdeq \tmext{A}{{P}{Q}}{\cprojfst{A}{\ctxext{P}{Q}}{w}}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}\\
& \jdeq \tmext{A}{{P}{Q}}{\cprojfst{A}{\ctxext{P}{Q}}{w}}{{\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}{\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}}\\
& \jdeq \tmext{{A}{P}}{Q}{{\cprojfst{A}{\ctxext{P}{Q}}{w}}{\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}}{\cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}
\end{align*}
Thus we see that 
\begin{align*}
\cprojfst{\ctxext{A}{P}}{Q}{w} & \jdeq \tmext{A}{P}{\cprojfst{A}{\ctxext{P}{Q}}{w}}{\cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}}}\\ 
\cprojsnd{\ctxext{A}{P}}{Q}{w} & \jdeq \cprojsnd{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
\end{align*}
proving the fourth judgmental equality, and therefore also that
\begin{align*}
\cprojfst{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}} & \jdeq \cprojfst{A}{\ctxext{P}{Q}}{w}\\
\cprojsnd{A}{P}{\cprojfst{\ctxext{A}{P}}{Q}{w}} & \jdeq \cprojfst{P}{Q}{\cprojsnd{A}{\ctxext{P}{Q}}{w}},
\end{align*}
proving the second and the third judgmental equalities.
\end{proof}

\begin{lem}
When we weaken a term $\tmext{B}{Q}{y}{v}$ of $\ctxext{B}{Q}$ in context $\Gamma$ by
a family $A$, the term that we get is $\tmext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A}{y}}{\ctxwk{A}{v}}$. More
precisely, the following inference rules are valid:
\begin{align*}
& \inference
  {\jterm{{\Gamma}{B}}{Q}{g}
   \jterm{{\Gamma}{B}}{\subst{g}{R}}{t}}
  {\jtermeq
    {{{\Gamma}{A}}{\ctxwk{A}{B}}}
    {\ctxwk{A}{\ctxext{Q}{R}}}
    {\ctxwk{A}{\tmext{Q}{R}{g}{t}}}
    {\tmext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}{\ctxwk{A}{g}}{\ctxwk{A}{t}}}}
% \\
%& \inference{\jterm{\Gamma}{\ctxwk{A}{\ctxext{B}{Q}}}{\ctxwk{A}{w}}}{\jtermeq{{\Gamma}{A}}{\ctxwk{A}{B}}{\cprojfst{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A}{w}}}{\ctxwk{A}{(\cprojfst{B}{Q}{w})}}}\\
\end{align*}
\end{lem}
